---
title: "Loading Raw and Processed Data"
author: "Keith Baggerly"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

I want to work through how to include raw and processed data files. 

Let's say we have a csv file, "basicData.csv", which we want to 
summarize. 

First, at the top level of the package (which has R/ and man/), we create a new folder for "inst/" (install) if this doesn't already exist. 

Next, within inst we similarly create a subfolder for "extdata".

This is where raw data should be put. 

We copy basicData.csv into inst/extdata.

Then, we recompile the package using "Ctrl-Shift-B". This will add the csv file as a "system file" for the package. 

Let's see where this is now in sisbid3.

```{r checkSisbid3}

system.file("extdata", "basicData.csv",
            package = "sisbid3")

```

This is the path to where the csv file resides on your system. Now we want to load this data, process it, and store the processed version for later use. 

```{r processTheData}

basicData <- 
  read.csv(system.file("extdata", "basicData.csv",
                       package = "sisbid3"))
basicData

```

Now we want to save this as a processed (.Rda) file. 

```{r saveTheData, eval=FALSE}

use_data(basicData)

```

This creates "data/basicData.rda". Now, as with all objects, we need to document the data. So, we go to R/, add "basicData.R":

```{r documentingBasicData, eval=FALSE}

#' A small basic dataset
#'
#' This is just an example to see if I can get
#' this to work
#'
#' @format A data frame with 5 rows and 2 columns
#' \describe{
#'   \item{basicDataX}{simulated X values}
#'   \item{basicDataY}{simulated Y values}
#' }
#' @source I made this up
"basicData"

```

Invoke document (to document the dataset), and recompile the package using Ctrl-Shift-B. At this point, the basicData dataset is available at any time. 

Now let's consider a more realistic example. 

Say we copy over the iomReview.csv to inst/extdata, and recompile the package so it's findable using system.file.

```{r showIomReview}

system.file("extdata", "iomReview.csv",
            package = "sisbid3")

```

Now we can load the contents into an R dataset called "vitD" and invoke "use_data".

```{r invokeVitD, eval=FALSE}

vitD <- 
  read.csv(system.file("extdata", 
                       "iomReview.csv", 
                       package = "sisbid3"))
vitD <- vitD[!is.na(vitD[,"N"]),]

use_data(vitD)

```

This creates "data/vitD.rda".

We still need to document it. Here, the full documentation is in the "vitD.R" file available from the Module3 website, so we simply copy this file over to R/.

Then we invoke document (to document the dataset), and then recompile the package. The vitD object is now part of the package.

